---
title: "27th september: Summarising Data"
author: "Tanya"
---

#### Introduction

Today we are learning to summarize data so we can get relevant data that is significant to us.

```{r}
#| label: setup
library(tidyverse)
library(mosaic)
library(skimr)
library(knitr)
library(kableExtra)
library(babynames)



```

Here we are taking a data set called mpg and and viewing the first 10 columns:

```{r}
mpg %>% 
  head(10)



```

-   qualitative data:

    -   manufacture

    -   model

    -   displacement

    -   year

    -   cyl

    -   trans

    -   drv

-   quantitative data:

    -   city millage

    -   high way millage

    #### Ways of Summarizing Data

```{r}
mpg %>% skimr::skim()

```

```{r}
glimpse(mpg)

```

```{r}
inspect(mpg)


```

Since, the computer R studio doesn't recognize that certain variables are Ordinal. Here is a code that converts the data to Ordinal (factor).

> you can see that *\<int\>* has changed to *\<fct\>*

```{r}
mpg_modified <- mpg %>%
  dplyr::mutate(
    cyl = as_factor(cyl),
    fl = as_factor(fl),
    drv = as_factor(drv),
    class = as_factor(class),
    trans = as_factor(trans)
  )
glimpse(mpg_modified)
```

#### Playing Around!!

I imported the library baby names and will try summarizing it.

```{r}
glimpse(babynames)
```

Now I will take **sex** and **factor** it because its an Ordinal data that has two levels.

> You can see that *\<char\>* has changed to *\<fct\>*

```{r}
babynames_modified <- babynames %>% 
  dplyr::mutate(
    sex= as.factor(sex)
  )
glimpse(babynames_modified)

```
