---
title: "10th October: Groups"
author: "Tanya"
---

### Groups

Today we will be looking at box plots and understanding the results we get from them. Box plots focuses on *median* and upper and lower *quartiles.*

```{r}
#| label: setup
library(tidyverse)
library(mosaic)
library(ggformula)
library(palmerpenguins)
```

#### Wages Data

```{r}
wages <- read_csv("https://vincentarelbundock.github.io/Rdatasets/csv/stevedata/gss_wages.csv")
```

```{r}
glimpse(wages)
inspect(wages)

```

```{r}
wages.mody<- wages %>% 
  mutate(
    gender=as_factor(gender),
    educcat=as_factor(educcat),
    maritalcat=as_factor(maritalcat),
    wrkstat=as_factor(wrkstat),
    occrecode=as_factor(occrecode)
    
  )
glimpse(wages.mody)
```

#### Data Munging

There is a lost of missing data in the variable *realrinc* so we will be dropping them.

```{r}
wages.mody <-
  wages.mody %>%
  tidyr::drop_na(realrinc) 
```

#### what is the basic distribution of realrinc?

```{r}

wages.mody %>%
  gf_boxplot( (realrinc)~ "Income") %>% # Dummy X-axis "variable"
  gf_labs(
    title = "Plot 1A: Income has a skewed distribution",
    subtitle = "Many outliers on the high side"
  )
```

#### Is Realrinc affected by gender?

```{r}
wages.mody %>%
  gf_boxplot(gender ~ realrinc) %>%
  gf_labs(title = "Plot 2A: Income by Gender")
```

```{r}
wages.mody %>%
  gf_boxplot(gender ~ log10(realrinc)) %>%
  gf_labs(title = "Plot 2B: Log(Income) by Gender")
```

```{r}
wages.mody %>%
  gf_boxplot(gender ~ realrinc, fill = ~gender) %>%
  gf_refine(scale_x_log10()) %>%
  gf_labs(title = "Plot 2C: Income filled by Gender, log scale")
```

#### Insights

-   Income is a highly skewed distribution

-   the mediums of male and female are highly varied.

-   the IQR is Lower for males than females.

-   there are outliers on both sides meaning that some people make a lots of money and some people make very less money.

#### Is Realinc affected by educcat?

```{r}
wages.mody %>%
  gf_boxplot(educcat ~ realrinc) %>%
  gf_labs(title = "Plot 3A: Income by Education Category")
```

```{r}
wages.mody %>%
  gf_boxplot(educcat ~ log10(realrinc)) %>%
  gf_labs(title = "Plot 3B: Log(Income) by Education Category")
```

```{r}
wages.mody %>%
  gf_boxplot(
    reorder(educcat, realrinc, FUN = median) ~ log(realrinc),
    fill = ~educcat,
    alpha = 0.3
  ) %>%
  gf_labs(title = "Plot 3C: Log(Income) by Education Category, sorted") %>%
  gf_labs(
    x = "Log Income",
    y = "Education Category"
  )
```

```{r}
wages.mody %>%
  gf_boxplot(reorder(educcat, realrinc, FUN = median) ~ realrinc,
    fill = ~educcat,
    alpha = 0.5
  ) %>%
  gf_refine(scale_x_log10()) %>%
  gf_labs(
    title = "Plot 3D: Income by Education Category, sorted",
    subtitle = "Log Income"
  ) %>%
  gf_labs(
    x = "Income",
    y = "Education Category"
  )
```

#### Insights

-   many Outliers therefor ppl do make very high or very low income

-   medians do increase in terms of educcat which means educcat does affect realrinc

#### Is Realrinc affected by combination of factors?

```{r}
wages.mody %>%
  drop_na() %>%
  gf_boxplot(reorder(educcat, realrinc) ~ log10(realrinc),
    fill = ~educcat,
    alpha = 0.5
  ) %>%
  gf_facet_wrap(vars(childs)) %>%
  gf_refine(scale_fill_brewer(type = "qual", palette = "Dark2")) %>%
  gf_labs(
    title = "Plot 4A: Log Income by Education Category and Family Size",
    x = "Log income",
    y = "No. of Children"
  )
```

```{r}
wages.mody %>%
  drop_na() %>%
  mutate(childs = as_factor(childs)) %>%
  gf_boxplot(childs ~ log10(realrinc),
    group = ~childs,
    fill = ~childs,
    alpha = 0.5
  ) %>%
  gf_facet_wrap(~gender) %>%
  gf_refine(scale_fill_brewer(type = "qual", palette = "Set3")) %>%
  gf_labs(
    title = "Plot 4B: Log Income by Gender and Family Size",
    x = "Log income",
    y = "No. of Children"
  )
```

#### Insights

-   realrinc increases with educcat regardless of family size till less than equal to 7 children

-   with gender for females the median slightly decreases while for males it increases till less than equal to 4 children.

-   males have a higher number of high income outliers which means they generally get a higher income.
