---
title: "4th Oct: Playing Around"
author: "Tanya"
---

### Playing Around

Here I'm going to take the wage and education data set and play around with it

```{r}
#| label: setup
library(tidyverse)
library(skimr)
library(ggformula)
library(crosstable)
library(mosaic)
```

```{r}
WE<- read_delim(file="../../data/SLID.csv",delim=",")
WE
```

```{r}
glimpse(WE)
```

```{r}
inspect(WE)
```

```{r}
WE.mody<- WE %>% 
  mutate(
    sex=as_factor(sex),
    language=as_factor(language)
  )
glimpse(WE.mody)
```

### Are the number of people of higher wages higher or lower than people with lower wages?

```{r}
gf_histogram(~wages,
  data = WE.mody,
  bins = 100,
  fill="orange",
  colour="black"
) %>%
  gf_labs(
    title = "Plot 1A: Wages",
    caption = "ggformula"
  )
```

> From this graph I can infer that people with the wages ranging 0-20 are significantly higher than 20-50.

```{r}
gf_histogram(~wages,
  data = WE.mody,
  bins = 100,
  fill= ~language,
  colour="black"
 ) %>%
  gf_labs(
    title = "Plot 2A: Wages and Languages",
    caption="ggformula"
  )
```

```{r}
gf_histogram(~wages,
  data = WE.mody,
  bins = 100,
  fill= ~language,
  colour="black",
  position="fill"
 ) %>%
  gf_labs(
    title = "Plot 2B: Wages and Languages",
    caption="ggformula"
  )

```

```{r}
WE.mody %>%
  gf_histogram(~wages, fill = ~language, color = "black", alpha = 0.3) %>%
  gf_facet_wrap(~language) %>%
  gf_labs(
    title = "Plot 2C: Wages Filled and Facetted by Languages",
    caption = "ggformula"
  ) %>%
  gf_theme(theme(
    axis.text.x = element_text(
      angle = 45,
      hjust = 1
    )
  ))


```

```{r}
WE.mody %>% 
  group_by(language) %>% 
  summarize(avg_w=mean(wages, na.rm=TRUE),count=n())
```

> We can infer from these graphs that English speaking people get more jobs but the average in all the language's wages are similar.

### Which sex receives a higher wage?

```{r}
gf_histogram(~wages,
  data = WE.mody,
  bins = 100,
  fill= ~sex,
  colour="black"
 ) %>%
  gf_labs(
    title = "Plot 2A: Wages and Sex",
    caption="ggformula"
  )
```

```{r}
gf_histogram(~wages,
  data = WE.mody,
  bins = 100,
  fill= ~sex,
  colour="black",
  position="fill"
 ) %>%
  gf_labs(
    title = "Plot 3B: Wages and Sex",
    caption="ggformula"
  )
```

```{r}
WE.mody %>% 
  group_by(sex) %>% 
  summarize(avg_w=mean(wages, na.rm=TRUE),count=n())
```

> > As we can infer from these graphs and table we can see that Males generally receive higher wages than Females.
